---
title: "Project 2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(data.table)){
  install.packages("data.table")
}

```

### Loading Data

Download and unpack files 

```{r data, cache = TRUE}
if (!file.exists("./tmp")) {
  dir.create("./tmp")
}

download.file("https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip", destfile = "./tmp/dataset.zip")

unzip("./tmp/dataset.zip",exdir = ".")
```

Observing unpacked directories

```{r}
dir("./UCI HAR Dataset")
dir("./UCI HAR Dataset/test")
dir("./UCI HAR Dataset/train")
```

### 1. Merging the training and the test sets to create one data set.

Observe datasets
```{r}
features <- fread("./UCI HAR Dataset/features.txt",sep = " ",header = FALSE, col.names = c("feature_id","feature"), stringsAsFactors = FALSE,encoding = "UTF-8")
dim(features) 
str(features)

activity_labels <- fread("./UCI HAR Dataset/activity_labels.txt", col.names = c("activity_id","activity_name"), stringsAsFactors = FALSE,encoding = "UTF-8")
dim(activity_labels) 
str(activity_labels)

```

Load and skim Train data
```{r}
x_train <- fread("./UCI HAR Dataset/train/X_train.txt", col.names = features$feature)
dim(x_train) 
str(x_train)

activity_labels <- fread("./UCI HAR Dataset/activity_labels.txt")
names(activity_labels) <- c("activity_id","activity_name")

y_train <- fread("./UCI HAR Dataset/train/y_train.txt")
names(y_train)[1] <- "activity_id"
dim(y_train) 
str(y_train)
```

Add Activities dimension 
```{r}
y_train$activity_factor <- factor(y_train$activity_id, levels=activity_labels$activity_id, labels = activity_labels$activity_name )
str(y_train)

if (dim(x_train)[1] == dim(y_train)[1]) {
  train_data <- cbind(y_train[,2],x_train)
  dim(train_data)
}
```

Load and observe Test data
```{r}
x_test<- fread("./UCI HAR Dataset/test/X_test.txt", col.names = features$feature)
dim(x_test) 
str(x_test)


y_test <- fread("./UCI HAR Dataset/test/y_test.txt")
names(y_test)[1] <- "activity_id"

y_test$activity_factor <- factor(y_test$activity_id, levels=activity_labels$activity_id, labels = activity_labels$activity_name )
str(y_test)

if (dim(x_test)[1] == dim(y_test)[1]) {
  test_data <- cbind(y_test[,2],x_test)
  dim(test_data)
}

```

Merge Training and Test datasets in one
```{r}
if (dim(test_data)[2] == dim(train_data)[2]) {
  all_data <- rbind(train_data,test_data)
  dim(all_data)
}
```

### 2. Extracting only the measurements on the mean and standard deviation for each measurement.

Using the RegExp filter *features* dataset, containing features list
```{r}

mean_std_cols <- grepl("std|mean",as.character(features$feature))
```

Remember that features was augmented by activity in 1st column, so include one mo TRUE value for it in the resulting logical vector/
```{r}
mean_std_cols <- c(TRUE,mean_std_cols)
mean_std_data <- all_data[,..mean_std_cols]

dim(mean_std_data)
str(mean_std_data)
```

### 3. Setting descriptive activity names to name the activities in the data set.
```{r}

```

### 4. Appropriately labeling the data set with descriptive variable names.
```{r}

```

### 5. From the data set in step 4, creating a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r}

```
